---
title: "Assignment 9"
author: "Rachel Greenlee"
date: "10/19/2020"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

For this assignment I'll be learning how to read data in from an API with the New York Times that requires a access key in order to look at data from their bestseller book lists the week of my birthday this past year. 



## Reading in the Data

First I'll need to install some packages to access, manipulate, and display the data

```{r, message=FALSE, warning=FALSE}
library(httr)
library(jsonlite)
library(dplyr)
library(stringr)
library(kableExtra)
```

Next I've put my API key for NYT in "nytkey" in a code chunk that is hidden, but I reference it below while using the GET function to access the API for NY Times Books.

```{r, echo = FALSE}
nytkey <- "tqMf5Luw5dXz238wNAUQrDiIhfkUgxZK"
```


```{r}
#path provided by NY Times 
path <- "https://api.nytimes.com/svc/books/v3/lists/overview.json"

#setting the date the bestseller list as my birthday, and a just-before the pandemic time
query <- "?published_date=2020-03-01"

#string it all together
url <- str_c(path, query, "&api-key=", nytkey, sep = "", collapse = NULL)

nyt_api <- GET(url, verbose())
```

## View List Types

Transform what we pulled in to a dataframe where we can look at the lists available to us.

```{r}
#grab the raw import of the API and set is text with proper encoding
bestseller_lists <- content(nyt_api, "text", encoding='UTF-8')

#convert from JSON
json_list <- fromJSON(bestseller_lists, flatten = TRUE)

#drill down to the lsits we want to see and display them
bestseller_lists <- json_list$results$list
kable(bestseller_lists$list_name, format = 'markdown')
```

## Accessing the Book Lists

Finally, lets look at two bestseller lists from my birthday week this year. First, I'll pull paperback nonfiction books and a description.

```{r}
#the 6 as it's the 6th list in the entry above
paperback_nf <- bestseller_lists[[6, "books"]]
paperback_nf <- subset(paperback_nf, select = c("title", "author", 
                                                "description", "weeks_on_list"))
kable(paperback_nf, format = 'pipe')
```


Next, lets look at the hardcover fiction books from my birthday week and pull slightly different variables.


```{r}
#the 1 as it's the 1st list in the entry above
hardcover_fic <- bestseller_lists[[1, "books"]]
hardcover_fic <- subset(hardcover_fic, select = c("title", "author", 
                                                  "publisher", 
                                                  "book_review_link"))
kable(hardcover_fic, format = 'pipe')
```


## Conclusion

While getting the URL correct to access the API took me some time, it was fairly intuitive to move through the file and pick out the best seller lists I wanted to see after that. 

For this week's assignment I had a hard time accessing the NY Times Books API, and found this vingette helpful in getting it to work:
https://cran.r-project.org/web/packages/jsonlite/vignettes/json-apis.html