---
title: "Project 1 - DATA607"
author: "Rachel Greenlee"
date: "9/14/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(stringr)
library(dplyr)
library(tidyr)
library(plyr)
```


# Step 1 - Import the text file  

I access the .txt file on my github repo. I skip the first 4 lines as it's dashes and headers. In order to to remove the full lines that comprise of dashes every 3 rows, I can write the pattern True, True, False for it to take the first and second rows, skip the third, and repeat.  

Now we read that into a table, with no header as we stripped it in line 1, and set the delimiter to a vertical bar/pipe. We tell it to fill in missing values in case the rows have unequal length for any player.

```{r}
raw.tourney <- read_lines("https://raw.githubusercontent.com/rachel-greenlee/data607_proj1/master/tournamentinfo.txt", 
                  skip = 4)[c(TRUE, TRUE, FALSE)]
raw.tourney <- read.table(textConnection(raw.tourney), header = FALSE, sep="|", fill = TRUE)
glimpse(raw.tourney)
```

\newpage
# Step 2 - Create data frame & clean it 

I create two data frames, one called first_rows and the other second_rows since this data has one chess players information across two rows. Then I add an ID, merge across on that ID, drop some unnecssary rows, and we have a full dataframe called d.tourney.

```{r}
#create 2 data frames
first_rows <- data.frame(first_rows <- raw.tourney %>%
 filter(row_number() %% 2 == 1))
second_rows <- data.frame(second_rows <- raw.tourney[c(rep(FALSE),TRUE),])

#create IDs for each data frame so we can match on it
first_rows$ID <- seq.int(nrow(first_rows))
second_rows$ID <- seq.int(nrow(first_rows))

#merge these two datasets together so all a player's data is in one row
d.tourney <- merge(first_rows, second_rows, by="ID")
#drop some columns we don't need
d.tourney <- subset(d.tourney, select = -c(V1.x, V11.x, V11.y))
```

Our dataframe columns need some cleaning so I create new columns and select the data I need.

```{r}
#create a pre-rating column and extract characters 15-18 to grab the relevant 
#part of the string
d.tourney$PreRating <- str_sub(d.tourney$V2.y, 15, 19)

#create a column for each of the possible 7 opponents, and start taking the 
#value from the existing column at character 2 as to avoid the letter 
#representing the outcome of the match, we just need the opponent's ID
#there are some blank entries at this point as not all players had 7 opponents
d.tourney$R1opp <- str_sub(d.tourney$V4.x, 2, )
d.tourney$R2opp <- str_sub(d.tourney$V5.x, 2, )
d.tourney$R3opp <- str_sub(d.tourney$V6.x, 2, )
d.tourney$R4opp <- str_sub(d.tourney$V7.x, 2, )
d.tourney$R5opp <- str_sub(d.tourney$V8.x, 2, )
d.tourney$R6opp <- str_sub(d.tourney$V9.x, 2, )
d.tourney$R7opp <- str_sub(d.tourney$V10.x, 2, )

#drop all those columns now that we have what we need from them
d.tourney <- subset(d.tourney, select = -c(V4.x, V5.x, V6.x, V7.x, V8.x, V9.x, 
                                           V10.x, V2.y, V3.y, V4.y, V5.y, V6.y, 
                                           V7.y, V8.y, V9.y, V10.y))
glimpse(d.tourney)
```

# Step 3 - Create calculated variables

Next I calculate the number of rounds each player had, out of a possible 7.

```{r}

#set the 7 variables of opponent IDs to be numeric and this also creates NAs in 
#the blank spaces
cols = c(6:12);    
d.tourney[,cols] = apply(d.tourney[,cols], 2, function(x) as.numeric(as.character(x)))

#sum up the number of NAs across each row, subtract for 7 possible games to get 
#the number of rounds that player plaid - store in "roundsplayed" variable
d.tourney$roundsplayed <- 7 - (apply(is.na(d.tourney), 1, sum)) 

```

Now we have to look-up the PreRating for all of a player's opponents, based on their ID so we can calculate the average PreRating of their opponents. I did this via a look-up table.

```{r}
#create a 2-variable look-up table with just the player ID and their PreRating
ratinglookup <- subset(d.tourney, select = c(ID, PreRating))

#we can overwrite the player ID that's in each RXopp variable with their 
#PreRating, which we can access by using the lookup table created above
d.tourney$R1opp = 
  ratinglookup[match(d.tourney$R1opp,
  ratinglookup$ID), "PreRating"]

d.tourney$R2opp = 
  ratinglookup[match(d.tourney$R2opp,
  ratinglookup$ID), "PreRating"]

d.tourney$R3opp = 
  ratinglookup[match(d.tourney$R3opp,
  ratinglookup$ID), "PreRating"]

d.tourney$R4opp = 
  ratinglookup[match(d.tourney$R4opp,
  ratinglookup$ID), "PreRating"]

d.tourney$R5opp = 
  ratinglookup[match(d.tourney$R5opp,
  ratinglookup$ID), "PreRating"]

d.tourney$R6opp = 
  ratinglookup[match(d.tourney$R6opp,
  ratinglookup$ID), "PreRating"]

d.tourney$R7opp = 
  ratinglookup[match(d.tourney$R7opp,
  ratinglookup$ID), "PreRating"]

#make a numeric class so we can do calculations
d.tourney <- d.tourney %>%
    mutate_at(vars(matches('R(.)opp')), list(as.numeric))

```

Finally, we create the variable that holds the average PreRating of the player's opponents.

```{r}
#sum the appropriate rows or the opponents' PreRatings then divide by the rounds 
#played
d.tourney$Avg_PreRating_of_Opponents = as.numeric(((rowSums(d.tourney[,c(6:12)], 
                                        na.rm = TRUE))/d.tourney$roundsplayed))

#round to nearest whole number
d.tourney$Avg_PreRating_of_Opponents <- round(d.tourney$Avg_PreRating_of_Opponents)

glimpse(d.tourney)
```

\newpage

# Step 4 - Final cleaning & export

Do some cleaning of the final dataframe.
```{r}
#remove unnecessary columns
d.tourney <- subset(d.tourney, select = -c(ID, R1opp, R2opp, R3opp, R4opp, 
                                           R5opp, R6opp, R7opp, roundsplayed))

#reorder to match assignment
d.tourney <- d.tourney[, c(1, 3, 2, 4, 5)]

#rename columns
d.tourney <- rename(d.tourney, c("V2.x"="Name", "V1.y"="State", 
                                 "V3.x"="TotalNumPoints"))

#capitlize names correctly
d.tourney$Name <- str_to_title(d.tourney$Name)

#trim whitespace around state abbreviations
d.tourney$State <- trimws(d.tourney$State, which = c("both"))

head(d.tourney)
```

Last, we write the dataframe to a CSV for export to the desktop. 

```{r}
write.csv(d.tourney,"C:\\Users\\rgreenlee\\Desktop\\\\chess_tournament_results.csv", 
          row.names = FALSE)
```

