---
title: "Hands on Lab: ggplot2"
author: "Catlin"
date: "September 15, 2015"
output: html_document
---

1. Load the diamonds dataset, which is included with the ggplot2 package.  Identify which variables in the diamond set are numeric, and which are categorical (factors).

```{r}
#install.packages("ggplot2")
require(ggplot2)
head(diamonds)

# Identify which variables in the diamonds dataset are numeric, and which
# variables are categorical
str(diamonds)
```

## Generate Descriptive Statistics

2. Generate summary level descriptive statistics:  Show the mean, median, 25th and 75th quartiles, min, and max for each of the applicable variables in diamonds.

See also [http://www.statmethods.net/stats/frequencies.html]

```{r}
summary(diamonds)
```

3. Determine the frequency for each of the diamond colors.
Run the table() function against a single categorical variable
See also: http://www.statmethods.net/stats/frequencies.html

```{r]}
table(diamonds$color)
```

4. Determine the frequency for each of the diamond cuts, by color.

```{r}
# Run the table() function against two categorical variables.
table(diamonds$color, diamonds$cut)
```

##R Graphics

5. Create a graph for a single numeric variable.

First, in base R.

```{r}
boxplot(diamonds$price)
hist(diamonds$price)
```

Look at the same single numeric variable, in ggplot2.

```{r}
# ggplot2:
qplot(price, data=diamonds)
```

6. Create a scatterplot of two numeric variables.

First, in base R:

```{r}
plot(diamonds$price ~ diamonds$carat)
```

Look at the same scatterplot in ggplot2.  Note that the order of carat and price are reversed.

```{r}
qplot(carat, price, data=diamonds)
```

